The Anatomy of a Temporal Summarization System

### 2015-01-24 algorithm/filtering.tex
edits to algo and problem definition mostly

\subsection{Filtering Content}
\label{sec:filtering}
We simulate our system using realistic data insofar as the document stream consists of relevant and non-relevant content.  As with many information retrieval systems, we crude, first phase of high-recall document filtering.  That is, we would like to remove most of the content which we can easily detect as unlikely to contain useful updates.

\subsubsection{Filtering Documents}
\label{sec:filtering:documents}
As a first filter, we retained only those documents whose raw html content contains at least one keyword from the event's query words. We further restricted our document set to only those articles from the news domain.\footnote{This filtered set was distributed by track organizers.}

\subsubsection{Filtering Sentences}
\label{sec:filtering:sentences}
In early versions of our summarization system, we found that structural html
artifacts and sentence tokenization errors
were negatively effecting the performance of later stages. Examples of the 
former include strings of link text like ``World Politics Sports ...'', while
examples of the latter included concatenations of various article headlines,
i.e. headlines pertaining to the event in question, as well as headlines from
other non-related events.
Both types of ``sentences'' were problematic for computing sentence similarity
as they were more likely to have higher average similarity to all input 
sentences. In turn they would be more likely to appear as cluster exemplars
in our clustering stage. From the clustering algorithm's point of view, this
is the correct decision to make---such multi-topic sentences are more general 
than a single topic, and better able to represent all aspects contained in a 
cluster. They make poor choices as updates, however, as they contain 
irrelevant information.

In order to filter out these problematic inputs, we trained a classifier to 
identify which sentences came from inside a document's main article and which
came from various headers, titles, menus, and links to other content. We collected ? random sentences and manually labeled whether the sentence came from
inside or outside the document's main article. We then trained a logistic 
regression classifier using the following features :

\begin{itemize}
 \item the position of the sentence
 \item word counts
 \item the last token in the sentence
 \item the last two tokens in the sentence
 \item the last three tokens in the sentence.
\end{itemize}

These features were sufficient to capture 
the main difference between content and non-content sentences,
which 
was that content sentences generally ended with sentence final punctuation, 
i.e. periods or a closing quotation mark.

Within our larger summarization framework, we process a document at a time,
identifying the subset of its sentences that are content sentences.
We then check to make sure \emph{all} event
query terms can be found within the document's content sentences. If so,
we send the content sentences on to the next stage of our pipeline; otherwise
we ignore all sentences in this document.

>>>>

\subsection{Filtering Content}
\label{sec:filtering}
We simulate our system using realistic data insofar as the document stream consists of relevant and non-relevant content.  As with many information retrieval systems, we use a crude, first phase of high-recall content filtering for both documents and sentences.

\subsubsection{Filtering Documents}
\label{sec:filtering:documents}
As a first filter, we retained only those documents whose raw html content contains at least one keyword from the event's query words. We further restricted our document set to only those articles from the news domain.\footnote{This filtered set was distributed by track organizers.} \fdcomment{not sure how much we have to describe this.}

\subsubsection{Filtering Sentences}
\label{sec:filtering:sentences}
Structural html artifacts and sentence tokenization errors can negatively effect the performance of a temporal summarization system. Examples of the former include strings of link text like ``World Politics Sports ...'', while examples of the latter included concatenations of unrelated content  (e.g. segmentation errors treating two headlines as a single sentence).

In order to filter out these problematic inputs, we trained a classifier to identify which sentences came from inside a document's main article and which came from various headers, titles, menus, and links to other content. We collected random sentences and manually labeled whether the sentence came from inside or outside the document's main article. We then trained a logistic  regression classifier using: 
\begin{enumerate*}[label=\itshape\alph*\upshape)]
	\item the position of the sentence,
	\item word counts,
	\item the last token in the sentence,
	\item the last two tokens in the sentence, and
	\item the last three tokens in the sentence.
\end{enumerate*}
These features were sufficient to capture the main difference between content and non-content sentences, which  was that content sentences generally ended with sentence final punctuation, i.e. periods or a closing quotation mark. \fdcomment{include summary performance numbers here; also a reference to prior work.}

After identifying the subset of a document's sentences that are content sentences, we check to make sure \emph{all} event query terms can be found within the document's content sentences. If so, we send the content sentences on to the next stage of our pipeline; otherwise we ignore all sentences in this document.

### 2015-01-24 algorithm/introduction.tex
edits to algo and problem definition mostly

We first begin with some notation. For a given event, let $\corpus$ be the set
of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$. 
Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let 
$\corpus_{\hour_i}$ be the set of retrieved documents such that 
$\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.

%Our TS system involves ? phases involving sentence level classification, 
%regression, and clustering. 
Figure ? outlines our general temporal summarization algorithm. The description
of our approach is as follows.
For each event, we iterate over the retrieved 
documents in hourly chunks, emitting 0 or more updates at each hour.
At each hour $\hour$, we process each document
$\doc \in \corpus_{\hour}$. First, we identify where the document content
actually is; Second, we predict the salience of all content sentences.
To account for redundancy, the predicted salience is penalized based on the 
distance of the sentence in question to the previous summary updates.

Finally, we cluster all content sentences 
for the current hour using the penalized salience predictions to bias the 
formation
of clusters around the most salient sentences. For each cluster center, or 
exemplar, that results, we check that the salience is above a threshold and 
that it does not belong to a singleton cluster; exemplars that satisfy these
conditions are emitted as an update.
Additionally, we maintain the complete set of updates in order to penalize
salience predictions in the subsequent time steps.

% \begin{figure}
% \centering 
\begin{algorithm}%[H]
 \KwData{$Query$ --- the set of event query words\\ 
        $SC$ --- the stream corpus\\

@@

 } 
 \caption{Temporal Summarization Algorithm}
\end{algorithm}
% \end{figure}

>>>>

Figure \ref{alg:temporal-summarization} outlines our general temporal summarization algorithm.  At each hour, the system processes each document in the batch. It first discards any documents below a predefined relevance threshold (Section \ref{sec:filtering:documents}).  It then discards any sentences from the remaining documents which are unlikely to carry content (Section \ref{sec:filtering:sentences}).  Following this filtering process, the system must decide which of the remaining sentences should be considered an update.  We refer to this as salience prediction and describe a feature-based model for predicting salience (Section \ref{subsec:Predict}).  Because a single sub-event may produce several high-quality-but-redundant sentences within an hour, we employ a cluster-based method for selecting exemplars of high salience sub-events (Section \ref{sec:ap}).  Finally, we employ a final deduplication phase to ensure that we only select novel updates (Section \ref{sec:novelty}).

% \begin{figure}
% \centering 
\begin{algorithm}%[H]
 \KwData{\\$Query$ --- the set of event query words\\ 
        $SC$ --- the stream corpus\\

@@

 } 
 \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
\end{algorithm}
% \end{figure}

### 2015-01-24 algorithm/salience.tex
edits to algo and problem definition mostly

\subsection{Predicting Sentence Salience}
\label{subsec:Predict}

In order to use AP clustering for summarization, we need to assign a
preference value to each input sentence.  In our approach, we equate a
sentence's salience with its preference.  A good model of sentence salience
should predict higher values for sentences that are more likely to appear in a
human generated summary of the event.

We do not have such human judgments, but we do have last year's gold nugget
sentences.
The response variable we try to predict is a sentence's semantic similarity
(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
to a gold nugget when the gold nugget is not known.

To build training data for this regression task, 
we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
sentences using the document retrieval and content selection steps outlined
in the previous selection. We then sample with replacement 200 sentences
from this collection, and extract non-lexical features (described in more 
detail below) for each sentence
to construct our design matrix $\mathbf{X}$. To build our response variables
$\mathbf{y}$ we compute the maximum semantic similarity of each input
sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
an RBF kernel to this data, optimizing kernel parameters with the scaled 
conjugate gradient method. This sampling procedure is repeated 100 times
for each of the ? 2013 TS events, yielding ? total models. In our salience 
predictions for the 2014 events, we take the mean prediction of all models.

%we take a subset of sentences
%relevant to the TREC events (approximately 1000) and match them to the gold
%nugget sentence with highest similarity as determined by the sentence
%similarity system of \cite{guo2012simple}. 
%\cite{} have used this system previously to correlate sentences to meaningful
%units of information in human generated summaries. 
%We use the real-valued similarity scores as our salience scores for the
%training sentences.

We want our model to be predictive across different kinds of events so we
avoid lexical features.  Instead, we extract a variety of features including
language model scores, geographic relevance, and temporal relevance from each
sentence.  These features are used to fit a Gaussian process regression model
that can predict the similarity of a sentence to a gold summary
\cite{preotiuc2013temporal}.  
%We use the model predicted salience of each
%sentence as it's preference value in the AP clustering.

\subsubsection{Features}
\paragraph{Basic Features}

We employ several basic features that have been used previously in supervised
models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
These include sentence length, the number of capitalized words normalized by
sentence length, document position, number of named entities, etc.

@@

These features are derived from the event query and the event type.
These include the number of query words present in the sentence
in addition to the number of event type synonyms, hypernyms, and hyponyms
as found in WordNet \cite{miller1995wordnet}.
E.g., for event type \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
``seism,'' ``aftershock,'' etc.

\paragraph{Language Model Features}\label{subsubsec:lm}

%Because the data in our experiments is scraped from the web, it is common to
%find sentences that contain both salient informantion and two kinds of noise:
%noisey fragments of web page structure (e.g. section titles, \emph{News},
%\emph{Sports}, etc.) and references to other news not relevant to the topic
%summary.
%
We use two trigram language models, trained using the SRILM toolkit
\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
the sentence's total log probability normalized by sentence length) from each
model.  This first model is trained on 4 years (2005-2009) of articles from
the Gigaword corpus.  Specifically, we use articles from the Associated Press
and the New York Times. This model is intended to assess the general writing
quality (grammaticality, word usage) of an input sentence and helps us to
filter out text snippets which are not sentences (e.g., web page titles).  The
second model is a domain specific language model. We build a corpus of
Wikipedia articles for each event type, consisting of documents from a related
Wikipedia category. E.g. for earthquakes, we collect pages under the category
\emph{Category:Earthquakes}. This model assigns higher probability to
sentences that are focused on the given domain.

@@

Locations are identified using a named entity tagger. For each location in a
sentence, we obtain its latitude and longitude using the Google Maps API.  We
then compute its distance to that of the event location.  It is possible for a
sentence and an event to have multiple locations so we take as features the
minimum, maximum, and average distance of all sentence-event location pairs.
Distances are calculated using the Vincenty distance.

@@

Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
D_{t-1}: w \in d^\prime\}|} \right) \end{align*}

@@

\subsubsection{Gaussian Process Regression}
\fdcomment{transition to gp description}

>>>>

\subsection{Predicting Sentence Salience}
\label{subsec:Predict}
Once the input has been filtered down to a higher quality set of candidate updates, we can rank sentences according to their salience.  We adopt a model-based approach to predicting salience: a good model should predict higher values for sentences that are more likely to appear in a human generated summary of the event.   The response variable we try to predict is a sentence's semantic similarity \cite{guo:wtmf} to the gold nugget sentences, i.e. we want to predict the similarity to a gold nugget when the gold nugget is not known.  If sentence is very similar to a nugget, then it is a better candidate for emission as an update compared to a sentence that is not similar to any nuggets.

\subsubsection{Features}
We want our model to be predictive across different kinds of events so we avoid lexical features.  Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.  These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.  
\fdcomment{can we enumerate all of these?}

\paragraph{Basic Features}

We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.

@@

Query features measure the relationship between the sentence and the event query and type.  These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.  For example, for event type \emph{earthquake},  we match sentence terms ``quake'', ``temblor'', ``seism'', and ``aftershock''.

\paragraph{Language Model Features}\label{subsubsec:lm}
Language models allow us to measure the likelihood of a sentence having been produced from a particular source.  We consider two types of language model features.  The first model is estimated from a corpus of generic news articles.  This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps us to filter out text snippets which are not sentences (e.g., web page titles).  The second model is estimated from text specific to our event types.  For example, the language model for event type `earthquake' is estimated from Wikipedia pages under the category \emph{Category:Earthquakes}.  These models are intended to detect sentences similar to those appearing in summaries of other events in the same category (e.g. most earthquake summaries are likely to include higher probability for ngrams including the token `magnitude').

@@

Locations are identified using a named entity tagger. For each location in a sentence, we obtain its latitude and longitude using the a publicly available geolocation service.  We then compute its distance to that of the event location.  It is possible for a sentence and an event to have multiple locations so we take as features the minimum, maximum, and average distance of all sentence-event location pairs.  Distances are calculated using the Vincenty distance.

@@

Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour burstiness of sentence $s$ from document $d$ at hour $t$  as 
\begin{align*}
\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
D_{t-1}: w \in d^\prime\}|} \right) \end{align*}

@@

\subsubsection{Gaussian Process Regression}

\fdcomment{transition to gp description}

### 2015-01-24 methods.tex
restructuring
edits to algo and problem definition mostly

\section{Methods}

\subsection{Data}\label{sec:data}

@@

@@

\begin{figure}
\includegraphics[scale=.40]{match-dist.png} 
\caption{Distribution of nugget similarity for matching and non-matching 
updates.}
\end{figure}
\subsubsection{Effect of Salience Models}

>>>>

\section{Methods}
\label{sec:methods}
\subsection{Data}\label{sec:data}

@@

\subsubsection{Estimating Language Models}
We use two trigram language models, trained using the SRILM toolkit \cite{stolcke2002srilm}, taking as features the average log probability (i.e. the sentence's total log probability normalized by sentence length) from each model.  This first model is trained on 4 years (2005-2009) of articles from the Gigaword corpus.  Specifically, we use articles from the Associated Press and the New York Times. The second model is a domain specific language model. We build a corpus of Wikipedia articles for each event type, consisting of documents from a related Wikipedia category. E.g. for earthquakes, we collect pages under the category \emph{Category:Earthquakes}. This model assigns higher probability to sentences that are focused on the given domain.

% For each event, we retrieve all sentences using the document retrieval and content selection steps outlined in the previous selection. We then sample with replacement 200 sentences from this collection, and extract non-lexical features for each sentence to construct our design matrix $\mathbf{X}$. To build our response variables $\mathbf{y}$ we compute the maximum semantic similarity of each input sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with an RBF kernel to this data, optimizing kernel parameters with the scaled  conjugate gradient method. This sampling procedure is repeated 100 times for each of the ? 2013 TS events, yielding \fdcomment{?} total models. In our salience predictions for the 2014 events, we take the mean prediction of all models.

@@

\begin{figure}
\includegraphics[scale=.40]{match-dist.png} 
\caption{Distribution of nugget similarity for matching and non-matching 
updates.\fdcomment{possible to output this as a vector graphic (pdf pref).  it will scale better.}}
\end{figure}
\subsubsection{Effect of Salience Models}

### 2015-01-24 paper.tex
restructuring
edits to algo and problem definition mostly

\documentclass{sig-alternate} 
\usepackage{url} 
\usepackage{color}
% \documentclass[10pt]{article} \usepackage{url} \usepackage{color}
\usepackage[utf8]{inputenc}
% \usepackage[margin=1in]{geometry}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[]{algorithm2e}
\DeclareMathOperator{\corpus}{\mathcal{C}}
\DeclareMathOperator{\doc}{\mathnormal{d}}
\DeclareMathOperator{\sent}{\mathnormal{s}}
\DeclareMathOperator{\order}{\pi}
\DeclareMathOperator{\dtime}{\mathnormal{t}}
\DeclareMathOperator{\hour}{\mathnormal{h}}
\DeclareMathOperator{\hours}{\mathcal{H}}
\DeclareMathOperator{\Sim}{\mathbf{K}}
\DeclareMathOperator{\SMat}{\mathbf{X}}
\DeclareMathOperator{\Pref}{\boldsymbol{\pi}}
\DeclareMathOperator{\Exemp}{\mathnormal{Exemplars}}
\DeclareMathOperator{\Updates}{\mathnormal{Updates}}
\usepackage{cleveref}
\crefname{section}{§}{§§}
\Crefname{section}{§}{§§}
\usepackage{graphicx}
\usepackage{authblk}
\newcommand{\fdadd}[1]{\textcolor{red}{#1}}
\newcommand{\fdcomment}[1]{\textbf{\textcolor{red}{[FD: #1]}}}
\newcommand{\ckcomment}[1]{\textbf{\textcolor{red}{[CK: #1]}}}

@@

\maketitle \begin{abstract} 
Information need is highest during times of crisis; both officials and 
civilians need effective access to make informed decisions about their 
safety and the safety of others. However, these situations pose many 
interesting challenges for the IR systems that support these decisions.
Specifically, we address event monitoring and updating, a.k.a. the
temporal summarization task (TS). Given an event query, we monitor streams
of online news and select important text to present to the user. Updates
are selected by predicting the salience of the text w.r.t. to the event.
Our model utilizes a variety of features at the query, document, and corpus
level. Additionally, these predictions are used to directly bias a clustering
algorithm for sentence selection, 
increasing the novelty of the updates. We evaluate our system
on a varied set of retrospective events using a combination of human 
judgements
and automatic evaluation measures, as compared to a baseline system.
We demonstrate the effect of the various feature groups, and the importance
of salience prediction w.r.t. update precision when compared to a clustering
only baseline.

@@

\input{introduction.tex}

@@

>>>>

\documentclass{sig-alternate} 
\usepackage{url} 
\usepackage{color}
% \documentclass[10pt]{article} \usepackage{url} \usepackage{color}
\usepackage[utf8]{inputenc}
% \usepackage[margin=1in]{geometry}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[inline]{enumitem}
\usepackage{amsfonts}
\usepackage[]{algorithm2e}
\usepackage{cleveref}
\crefname{section}{§}{§§}
\Crefname{section}{§}{§§}
\usepackage{graphicx}
\usepackage{authblk}
\newcommand{\fdadd}[1]{\textcolor{red}{#1}}
\newcommand{\fdcomment}[1]{\textbf{\textcolor{red}{[FD: #1]}}}
\newcommand{\ckcomment}[1]{\textbf{\textcolor{blue}{[CK: #1]}}}
\newcommand{\kmcomment}[1]{\textbf{\textcolor{green}{[KM: #1]}}}
\input{notation.tex}

@@

\maketitle \begin{abstract} 
During crises such as natural disasters or other human tragedies, information needs of both civilians and responders often require urgent, specialized treatment.  Slow or ineffective information access can significantly impact the safety and health of individuals.  Monitoring and summarizing a text stream during such an event remains a difficult problem.  The temporal summarization task refers to extracting short, sentence-length updates about a seed event from a stream of text data.  We present a system which makes substantial improvements over the state of the art in temporal summarization.  Our approach predicts the salience of sentences with respect to an event and then uses these predictions to directly bias a clustering algorithm for sentence selection, increasing the novelty of the updates. We use novel, disaster-specific features for salience prediction, including geo-locations and language models representing the language typically used to describe different disaster types. 
We evaluate our system on a standard set of retrospective events using a combination of human judgements and automatic evaluation measures. We demonstrate the effect of different feature groups, and the importance of salience prediction for update precision.

@@

\fdcomment{We should try to pull out as much of the TREC-specific language as possible (e.g. `run submissions', `participation in the TS track') to make the results more general.}
\input{introduction.tex}

@@

\input{problemdefinition.tex}

### 2015-01-24 problemdefinition.tex
edits to algo and problem definition mostly

A temporal summarization system takes as input 
\begin{enumerate*}[label=\itshape\alph*\upshape)]
  \item a short query $\query$ defining the event to be tracked (e.g. `Hurricane Sandy'), 
  \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$, presented in temporal order, and
	\item an evaluation time period of interest, $(\stime,\etime)$.  
\end{enumerate*}
Throughout the time period of interest, the system must output sentences, known as \emph{updates}, which are \emph{relevant} to the query, \emph{comprehensive} with respect to the event, \emph{novel} to the user, and \emph{timely} with respect to when the update occurred.  We present the evaluation metrics which measure these properties in Section \ref{sec:methods}.

In the ideal version of this problem, the system makes decisions every time it consumes a document $\doc_i$.  In practice, we assume that that we receive batches of new sentence-segmented documents every hour.

>>>>

A temporal summarization system takes as input 
\begin{enumerate*}[label=\itshape\alph*\upshape)]
  \item a short query $\query$ defining the event to be tracked (e.g. `Hurricane Sandy'), 
  \item an event category $\eventcategory$ defining the type of event to be tracked (e.g. `hurricane'), 
  \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$, presented in temporal order, and
	\item an evaluation time period of interest, $(\stime,\etime)$.  
\end{enumerate*}
While processing documents throughout the time period of interest, the system must output sentences, known as updates, which are \emph{relevant} to the query, \emph{comprehensive} with respect to the event, \emph{novel} to the user, and \emph{timely} with respect to when the update occurred.  Precisely how to measure these properties will be discussed in Section \ref{sec:methods}.  In our version of this problem, we assume that that the system receives one batch of new sentence-segmented documents every hour throughout the period of interest.

\fdcomment{include description of nuggets here.}
